---
title: "Importing and Managing Financial Data in R"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
# Data importing
## getSymbols
Search for symbols at the sources websites

Sources:
- Yahoo! Finance
- Google Finance
- FRED
- Oanda
- CSV

```{r}
library(quantmod)
symbls = c("AAPL")
getSymbols(symbls, auto.assign = T)
head(AAPL, n=2)
str(AAPL)
```

### Setting getSymbols default arguments

```{r}
# do not persist across R sessions
setDefaults(getSymbols, src="yahoo")
getDefaults()
getDefaults(getSymbols)
args(getSymbols.yahoo)
```

### Syntactically invalid names

```{r}
# getSymbols("000001.SS", auto.assign = T)
# str(000001.SS)

# getSymbols(`000001.SS`, auto.assign = T)
# str(`000001.SS`)

SSE <- getSymbols("000001.SS", auto.assign = F)
str(SSE)

setSymbolLookup(SSE = list(name = "000001.SS"))
getSymbols("SSE", auto.assign = T)
str(SSE)

saveSymbolLookup("my_symbl_lookup")
loadSymbolLookup("my_symbl_lookup")
```

```{r}
# Create BRK.A object
BRK.A = getSymbols("BRK-A", auto.assign=F)

# Create col_names object with the column names of BRK.A
col_names <- colnames(BRK.A)

# Set BRK.A column names to syntactically valid names
colnames(BRK.A) <- make.names(col_names, unique=T)
```

## Quandl

```{r}
# Load the Quandl package
# library(Quandl)
# 
# # Import GDP data from FRED
# gdp <- Quandl("FRED/GDP")
# 
# # Look at the structure of the object returned by Quandl
# str(gdp)
# 
# 
# gdp <- Quandl("FRED/GDP", type = "xts")
# str(gdp)
```

## Google Finance

```{r}
# # Create an object containing the Pfizer ticker symbol
# symbol <- "AAPL"
# 
# # Use getSymbols to import the data
# getSymbols(symbol, src="google", auto.assign=T)
# 
# 
# # Look at the first few rows of data
# head(PFE)
```

## Oanda

```{r}
# Create a currency_pair object
currency_pair <- "GBP/CAD"

# Load British Pound to Canadian Dollar exchange rate data
getSymbols(currency_pair, auto.assign=T, src="oanda")

# Examine object using str()
str(GBPCAD)

# Try to load data from 190 days ago
getSymbols(currency_pair, from = Sys.Date() - 190, to = Sys.Date(), src = "oanda")
```

## FRED

```{r}
# Create a series_name object
series_name <- "UNRATE"

# Load the data using getSymbols
getSymbols(series_name, src="FRED")

# Create a quandl_code object
# quandl_code <- "FRED/UNRATE"

# Load the data using Quandl
# unemploy_rate <- Quandl(quandl_code)
```

## Importing text files

```{r}
?read.zoo()
# # Read data with read.csv
# une_data <- read.csv("UNE.csv", nrows = 5)
# 
# # Look at the structure of une_data
# str(une_data)
# 
# # Read data with read.zoo, specifying index columns
# une_zoo <- read.zoo("UNE.csv", index.column = c("Date", "Time"), sep = ",", header = T)
# 
# # Look at first few rows of data
# head(une_zoo)
```



# Data manipulation
## Extracting Columns

```{r}
head(AAPL, n=2)
aapl_close <- Cl(AAPL)
head(aapl_close, n=2)

aapl_volume <- Vo(AAPL)
head(aapl_volume, n=2)

aapl_ohlc <- OHLC(AAPL)
head(aapl_ohlc, n=2)

```

```{r}
# Download CME data for CL and BZ as an xts object
oil_data <- Quandl(code = c("CME/CLH2016", "CME/BZH2016"), type = "xts")

# Look at the column names of the oil_data object
colnames(oil_data)

# Extract the Open price for CLH2016
cl_open <- getPrice(oil_data, symbol = "CLH2016", prefer = "Open$")

# Look at January, 2016 using xts' ISO-8601 subsetting
cl_open["2016-01"]
```

## Multiple Instruments (environment)

```{r}
data_env <- new.env()
getSymbols(c("SPY", "QQQ"), env = data_env, auto.assign = T)
data_env
head(data_env$SPY, n=2)


adjusted_list <- lapply(data_env, Ad)
adjusted <- do.call(merge,adjusted_list)
head(adjusted, n=2)
```

## Regularising irregular data

```{r}
from_date <- as.Date("2017-01-01")
to_date <- as.Date("2017-12-01")
months <- seq(from_date,
              to_date, 
              by = "month")
head(months, n=3)

regular_xts <- xts(seq_along(months), 
                   order.by = months)
start(regular_xts)
end(regular_xts)

days <- seq(from = start(regular_xts), 
            to = end(regular_xts), 
            by = "day")

# zero-width xts object
zero_width_xts <- xts(,order.by = months)
zero_width_xts

irregular <- xts(rnorm(5), order.by = from_date+seq(0,300, length.out = 5))
irregular

date_seq <- seq(from = start(irregular), 
            to = end(irregular), 
            by = "month")

new_regular_xts <- xts(,order.by = date_seq)
new_regular_xts <- merge(irregular, new_regular_xts)
head(new_regular_xts, h=5)
```

## Timezones

```{r}
# # Create merged object with a Europe/London timezone
# tz_london <- merge(london, chicago)
# 
# # Look at tz_london structure
# str(tz_london)
# 
# # Create merged object with a America/Chicago timezone
# tz_chicago <- merge(chicago, london)
# 
# # Look at tz_chicago structure
# str(tz_chicago)
```

```{r}
# # Create a regular date-time sequence
# regular_index <- seq(as.POSIXct("2010-01-04 09:00"), as.POSIXct("2010-01-08 16:00"), by = "30 min")
# 
# # Create a zero-width xts object
# regular_xts <- xts(order.by=regular_index)
# 
# # Merge irregular_xts and regular_xts, filling NA with their previous value
# merged_xts <- merge(irregular_xts, regular_xts, fill=na.locf)
# 
# # Subset to trading day (9AM - 4PM)
# trade_day <- merged_xts['T09:00/T16:00']
```

```{r}
# # Split trade_day into days
# daily_list <- split(trade_day , f = "days")
# 
# # Use lapply to call na.locf for each day in daily_list
# daily_filled <- lapply(daily_list, FUN = na.locf)
# 
# # Use do.call to rbind the results
# filled_by_trade_day <- do.call(rbind, daily_filled)
```

```{r}
# # Convert raw prices to 5-second prices
# xts_5sec <- to.period(intraday_xts, period = "seconds", k = 5)
# 
# # Convert raw prices to 10-minute prices
# xts_10min <- to.period(intraday_xts, period = "minutes", k=10)
# 
# # Convert raw prices to 1-hour prices
# xts_1hour <- to.period(intraday_xts, period = "hours", k = 1 )
```

## Weirdness
- Missing values (see xts-zoo notebook)
- Stock splits and dividends
```{r}
getSymbols("MSFT", auto.assign = T)
msft_adjusted <- adjustOHLC(MSFT)

?adjRatios
```

```{r}
# Download AAPL split data
splits <- getSplits("AAPL")

# Print the splits object
splits

# Download AAPL dividend data
dividends <- getDividends("AAPL")

# Look at the first few rows of dividends
head(dividends)

# Download unadjusted AAPL dividend data
raw_dividends = getDividends("AAPL", split.adjust = F)

# Look at the first few rows of raw_dividends
head(raw_dividends)
```

```{r}
# Calculate split and dividend adjustment ratios
ratios <- adjRatios(splits = splits, dividends = raw_dividends, close = Cl(AAPL))

# Calculate adjusted close for AAPL
aapl_adjusted <- Cl(AAPL) * ratios[, "Split"] * ratios[, "Div"]

# Look at first few rows of Yahoo adjusted close
head(Ad(AAPL))

# Look at first few rows of aapl_adjusted
head(aapl_adjusted)
```

