---
title: "Intermediate Portfolio Analysis in R"
output:
  html_document:
    df_print: paged
---
### Mean - Standard Deviation Setup

```{r}
library(PortfolioAnalytics)
data(indexes)
index_returns <- indexes[,1:5]
```

```{r eval=FALSE, include=FALSE}
port_spec <- portfolio.spec(colnames(index_returns))

# add objectives
port_spec <- add.objective(port_spec,
                           type = "return",
                           name = "mean")
port_spec <- add.objective(port_spec,
                           type = "risk",
                           name = "StdDev")

# add constraints
port_spec <- add.constraint(port_spec, 
                            type = "full_investment")
port_spec <- add.constraint(port_spec, 
                            type = "long_only")

print(port_spec)
```

Run Optimization and Chart Results

```{r eval=FALSE, include=FALSE}
opt <- optimize.portfolio(index_returns, port_spec, optimize_method = "random", trace = T)
```
```{r eval=FALSE, include=FALSE}
chart.RiskReward(opt, risk.col = "StdDev", return.col = "mean", chart.assets = T)

# Print the results of the optimization
print(opt)

# Extract the optimal weights
extractWeights(opt)

# Chart the optimal weights
chart.Weights(opt)
```


### Solvers

 - Close solvers (quadreatic programming) -> FAST / EXACT
 - Global solvers (differential evolution optimization) -> SLOWER / APPROXIATE
 
### Quadratic Utility

Maximise: $\mathbf{w}' \mathbf{\mu} - \lambda * \mathbf{w}' \mathbf{\Sigma}\mathbf{w}$
Subject to:
- $w_i \ge 0$
- $\sum_{i=1}^N w_i = 1$

Where:
- $\mathbf{w}$ is the weight vector
- $\mathbf{\mu}$ is the expected return vector
- $\lambda$ is the risk aversion parameter
- $\mathbf{\Sigma}$ is the variance - covariance matrix

### Quadratic Programming Solver

package **quadprog**

**solve.QP()** solves problems in the form
- $min(-\mathbf{d}'\mathbf{b}+\frac{1}{2}\mathbf{b}'\mathbf{D}\mathbf{b})$

subject to:
- $\mathbf{A}'\mathbf{b} \ge b_0$
 
```{r eval=FALSE, include=FALSE}
library(quadprog)
# Constraint matrix
Amat <- cbind(1, diag(ncol(index_returns)), -diag(ncol(index_returns)))

# Constraint vector
bvec <- c(1, rep(0, ncol(index_returns)), -rep(1, ncol(index_returns)))

# Objective matrix
Dmat <- 10 * cov(index_returns)

# Objective vector
dvec <- colMeans(index_returns)

# Specify number of equality constraints
meq <- 1

# Solve the optimization problem
opt <- solve.QP(Dmat, dvec, Amat, bvec, meq)
opt
```


## Portfolio specs, constraints and objectives

```{r}
?portfolio.spec
?add.constraint
?add.objective
```

```{r include=FALSE}
# Create the portfolio specification
port_spec <- portfolio.spec(assets = colnames(index_returns))

# Add a full investment constraint such that the weights sum to 1
port_spec <- add.constraint(portfolio = port_spec, type = "full_investment")

# Add a long only constraint such that the weight of an asset is between 0 and 1
# port_spec <- add.constraint(portfolio = port_spec, type = "long_only")

# Add weight constraint
port_spec <- add.constraint(portfolio = port_spec, type = "box", min = 0.01, max = 0.8)

# Add an objective to maximize portfolio mean return
port_spec <- add.objective(portfolio = port_spec, type = "return", name = "mean")

# Add an objective to minimize portfolio variance
# port_spec <- add.objective(portfolio = port_spec, type = "risk", name = "var", risk_aversion = 10)
port_spec <- add.objective(portfolio = port_spec, type = "risk", name = "StdDev")
```

## Optimization Methods

```{r}
# Solve the optimization problem
opt_single <- optimize.portfolio(R = index_returns, portfolio = port_spec, optimize_method = "random", trace = T, search_size = 1000)

# Run the optimization backtest with quarterly rebalancing
opt_rebal <- optimize.portfolio.rebalancing(R = index_returns, portfolio = port_spec, optimize_method = "random", trace = T, search_size = 1000, rebalance_on = "quarters", training_period = 60, rolling_window = 60)
```

## Visualisation

```{r}
chart.RiskReward(opt_single, risk.col = "StdDev", return.col = "mean", chart.assets = T)

# Print the results of the optimization
print(opt_single)

# Extract the objective measures for the single period optimization
extractObjectiveMeasures(opt_single)

# Extract the optimal weights
extractWeights(opt_single)

# Chart the optimal weights
chart.Weights(opt_single)
```

```{r}
#chart.RiskReward(opt_rebal, risk.col = "StdDev", return.col = "mean", chart.assets = T)

# Print the results of the optimization
print(opt_rebal)

# Extract the objective measures for the optimization backtest
extractObjectiveMeasures(opt_rebal)

# Extract the optimal weights
extractWeights(opt_rebal)

# Chart the optimal weights
chart.Weights(opt_rebal)
```

## Analysis

```{r}
rr_single <- Return.portfolio(index_returns, weights = extractWeights(opt_single))
charts.PerformanceSummary(rr_single)
```

```{r}
rr_rebal <- Return.portfolio(index_returns, weights = extractWeights(opt_rebal))
charts.PerformanceSummary(rr_rebal)
```

## Moments

- 1st moment: Expected Returns Vector
- 2nd moment: variance-covariance matrix
- 3rd moment: coskewness matrix
- 4th moment: cokurtosis matrix

```{r}
# Sample vs Boudt (Factor)
sample_moments <- set.portfolio.moments(R = index_returns, portfolio = port_spec)
sample_moments$sigma == cov(index_returns)

fit <- statistical.factor.model(R = asset_returns, k = 1)

boudt_moments <- set.portfolio.moments(R = index_returns, portfolio = port_spec, method = "boudt", k=1)
boudt_moments$sigma == extractCovariance(fit)

```

### Custom Moment Fucntions

```{r eval=FALSE, include=FALSE}
library(MASS)
custom_fun <- function(R, portfolio, rob_method = "mcd") {
  out <- list()
  out$sigma <- cov.rob(R,method = rob_method)
  return(out)
}

port_spec1 <- portfolio.spec(colnames(index_returns))
port_spec1 <- add.constraint(portfolio = port_spec1, type = "full_investment")
port_spec1 <- add.constraint(portfolio = port_spec1, type = "long_only")

port_spec1 <- add.objective(portfolio = port_spec1, type = "risk", name = "StdDev")

opt1 <- optimize.portfolio(index_returns, port_spec1, optimize_method = "random", momentFUN = "custom_fun", rob_method = "mcd")

opt2 <- optimize.portfolio(index_returns, port_spec1, optimize_method = "random",  momentFUN = "custom_fun", rob_method = "mve")
```

```{r eval=FALSE, include=FALSE}
# Print the results of the optimization
print(opt1)

# Extract the objective measures for the optimization backtest
extractObjectiveMeasures(opt1)

# Extract the optimal weights
extractWeights(opt1)

# Chart the optimal weights
chart.Weights(opt1)

rr1 <- Return.portfolio(index_returns, weights = extractWeights(opt1))
charts.PerformanceSummary(rr1)
```

```{r eval=FALSE, include=FALSE}
# Print the results of the optimization
print(opt2)

# Extract the objective measures for the optimization backtest
extractObjectiveMeasures(opt2)

# Extract the optimal weights
extractWeights(opt2)

# Chart the optimal weights
chart.Weights(opt2)

rr2 <- Return.portfolio(index_returns, weights = extractWeights(opt2))
charts.PerformanceSummary(rr2)
```

## Objective Fucntions

```{r eval=FALSE, include=FALSE}
sr_annualised <- function(R, weights, sigma, scale, rfr) {
  r <- Return.annualized.excess(Return.portfolio(R, weights), scale = scale, Rb = rfr)
  
  pasd <- sqrt(as,numeric(t(weights) %*% sigma %*% weigths)) * sqrt(scale)
  return(r/pasd)
}


# Add custom objective to portfolio specification
port_spec <- add.objective(portfolio = port_spec, type = "return", name = "sr_annualised")
```

# Example
## Benchmark

```{r}
# Load the package
library(PortfolioAnalytics)

# Load the data
data(edhec)

# Assign the data to a variable
asset_returns <- edhec

# Create a vector of equal weights
equal_weights <- rep(1 / ncol(asset_returns), ncol(asset_returns))

# Compute the benchmark returns
r_benchmark <- Return.portfolio(R = asset_returns, weights = equal_weights, rebalance_on = "quarters")
colnames(r_benchmark) <- "benchmark"

# Plot the benchmark returns
plot.zoo(r_benchmark)
charts.PerformanceSummary(r_benchmark)
```
## Portfolio Spec

```{r}

# Create the portfolio specification
port_spec <- portfolio.spec(assets = colnames(asset_returns))

# Add a full investment constraint such that the weights sum to 1
port_spec <- add.constraint(portfolio = port_spec, type = "full_investment")

# Add a long only constraint such that the weight of an asset is between 0 and 1
port_spec <- add.constraint(portfolio = port_spec, type = "long_only")

# Add an objective to minimize portfolio standard deviation
port_spec <- add.objective(portfolio = port_spec, type = "risk", name = "StdDev")

# Print the portfolio specification
print(port_spec)
```

## Basic Optimization

```{r}
# Run the optimization
opt_rebal_base <- optimize.portfolio.rebalancing(R = asset_returns, portfolio = port_spec, optimize_method = "random", trace = T, search_size = 1000, rebalance_on = "quarters", training_period = 60, rolling_window = 60)

# Print the results
print(opt_rebal_base)

# Chart the weights
chart.Weights(opt_rebal_base)

# Compute the portfolio returns
returns_base <- Return.portfolio(R = asset_returns, weights = extractWeights(opt_rebal_base))
colnames(returns_base) <- "base"
```

```{r}
port_spec_rb <- port_spec
```

