---
title: "xts/zoo Notebook"
output:
  html_document:
    df_print: paged
---

eXtensible Time Series
Matrix + Index

```{r}
library(xts)
# Create the object data using 5 random numbers
data = rnorm(5)

# Create dates as a Date class object starting from 2016-01-01
dates <- seq(as.Date("2016-01-01"), length = 5, by = "days")

# Use xts() to create smith
smith <- xts(x = data, order.by = dates)

# Create bday (1899-05-08) using a POSIXct date class object
bday <- as.POSIXct("1899-05-08")

# Create hayek and add a new attribute called born
hayek <- xts(x = data, order.by = dates, born = bday)
```
Core Data, Index and Attributes within the xts object

```{r}
# Extract the core data of hayek
hayek_core = coredata(hayek)
hayek_core
# View the class of hayek_core
class(hayek_core)

# Extract the index of hayek
hayek_index = index(hayek)
hayek_index
# View the class of hayek_index
class(hayek_index)

xtsAttributes(hayek)
```

## Importing, Exporting and Converting

```{r}
## Importing
# as.xts(read.table("filename"))
# as.xts(read.csv("filename"))
# as.xts(read.zoo("filename"))

## Exporting
# write.zoo(x, "filename", sep=',', col.names = c("Date", "Sunspots"))
# saveRDS(x, "filename")

## Converting
data(sunspots)
class(sunspots)

sunspots_xts <- as.xts(sunspots)
class(sunspots_xts)

head(sunspots_xts, n=2)
```

Importing Example

```{r}
# Create dat by reading tmp_file
dat = read.csv("tmp.csv")
head(dat)

# Convert dat into xts
dat_xts <- xts(dat[-1], order.by = as.yearmon(dat[,1], "%b %Y"))

# Read tmp_file using read.zoo
f <- function(x) as.yearmon(x, "%b %Y")
dat_zoo <- read.zoo("tmp.csv", index.column = 1, sep = ",", header=T, FUN = f)

# Convert dat_zoo to xts
dat_xts_2 <- as.xts(dat_zoo)
```

## Time based queries
ISO 8601:2004

- One and two sided intervals
"2004"
"2001/2005"

- Truncated Representation
"201402/03"

- Time support
"2014-02-22 08:30:00"

- Repeating intervals
"T08:00/T09:00"

```{r}
# Select all of 1800 from x
x_1800 <- dat_xts_2["1800/1801"]

# Select January 1, 1800 to March, 1800
jan_march <- dat_xts_2["1800/1800-03"]

# Verify that jan_march contains 3 rows
3 == length(jan_march)

index <- dat_xts_2["1800/1800-03", which.i = T]
index
```

## First and Last

```{r}
# Create lastweek using the last 3 months of dat_xts_2
lastthreemonths <- last(dat_xts_2, "3 months")

# Print the first 2 observations in lastthreemonths
first(lastthreemonths, "2 months")
```

## Adding based on date vs adding to all

```{r}
head(dat_xts_2, n=5)
dat_first_2 = first(dat_xts_2, "2 months")
new_dat <- dat_xts_2 + dat_first_2
head(new_dat, n=5)

new_data_plus_5 <- dat_xts_2 + 5
head(new_data_plus_5, n=5)

```

## More adding

```{r}
# Add a to b, and fill all missing rows of b with 0
new_fill_0 <- dat_xts_2 + merge(dat_first_2, index(dat_xts_2), fill = 0)
head(new_fill_0, n=5)

# Add a to b and fill NAs with the last observation
new_fill_locf <- dat_xts_2 + merge(dat_first_2, index(dat_xts_2), fill = na.locf)
head(new_fill_locf, n=5)

```

## Merging Columns

```{r}
a <- xts(1:4, order.by = as.Date("2018-01-01")+0:3)
b <- xts(10:13, order.by = as.Date("2018-01-01")+2:5)
merge(a, b)  #outer
merge(a, b, join = "inner")
merge(a, b, join = "left", fill = 0)
merge(a, b, join = "right", fill = 0)
```

## Merging Rows

```{r}
rbind(a, b)
```

## Missing Values

```{r}
# na.locf(a, na.rm = T, fromLast = F)
z <- xts(c(1, NA, NA, 4, NA), order.by = as.Date("2018-01-01")+0:4)
cbind(z, na.locf(z), na.locf(z, fromLast = T))

na.fill(z, 10)

na.trim(z)

na.omit(z)
na.approx(z)
```

## Lag and Diff operations

```{r}
lag_and_diff <- merge(a, 
      lag(a,k=1),
      lag(a,k=2), 
      diff(a,lag= 1 ,differences = 1),
      diff(a,lag= 2 ,differences = 1))
colnames(lag_and_diff) <- c("a", "1 lag", "2 lag", "1 lag diff", "2 lag diff")
lag_and_diff
```

## Applying on time

```{r}
c <- xts(rnorm(500), order.by = as.Date("2018-01-01")+0:499)
# Locate the months
ep <- endpoints(c, on = "months")

# Locate every two months
ep2 <- endpoints(c, on = "months", k = 2)

# METHOD 1
# Calculate the monthly average
monthly_average <- period.apply(c, INDEX = ep, FUN = "mean")

# METHOD 2
# Split c by month
c_monthly <- split(c, f = "months")

# Create a list of weekly means, temps_avg, and print this list
c_avg <- lapply(X = c_monthly, FUN = mean)
monthly_average_2 <- do.call(rbind, c_avg)
```

## Periods

```{r}
c_weekly <- to.period(c, period = "weeks")
c_monthly <- to.period(c, period = "months")
c_quarterly <- to.period(c, period = "quarters")
c_yearly <- to.period(c, period = "years")
periodicity(c)
ndays(c)
nyears(c)
periodicity(c_yearly)
nyears(c_yearly)
.index(c_yearly)
.indexwday(c_yearly)

# ?make.index.unique
```

## Rolling

```{r}
# splits
c_yrs <- split(c, f="years")
c_yrs <- lapply(c_yrs, cumsum)
c_ytd <- do.call(rbind, c_yrs)

# windows
c_rolling_average <- rollapply(c, width = 25, FUN = mean)
```

